---
title: "CalibreMD"
author: "Ken Brooks"
date: "5/11/2025"
output: html_document
---

# Setup

```{r setup, include=FALSE}
library(CalibreMD)
setup_packages()

# Set knitr options directly
knitr::opts_chunk$set(
  fig.pos = 'h',
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  autodep = TRUE,
  cache = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  fig.show = "hold"
)

dataDir <- '/Users/kbrooks/Dropbox/Books/Calibre Travel Library'
#dataDir <- '/Users/kbrooks/Dropbox/Books/AI Travel Library'

md_db <- find_md_db(dataDir)
```

# Load data from SQLite to EAV table

```{r}
eav <- md_db %>% 
  load_eav() %>% 
  explode_text_features()
```

## Explore data
```{r}
tag_counts <- eav %>% get_tag_counts()
book_summary <- eav %>% get_book_summary()
tag_summary <- eav %>% get_tag_summary()
```

# Train model

```{r}
dataset <- prep_dataset(eav)
models <- train_models(dataset)
```

# Predict on all data

```{r}
predictions_probs <- predict_tags(models, dataset)
predictions_discrete <- threshold_predictions(predictions_probs, book_id = 900, threshold = 0.9)
top_n_labels <- get_top_n_labels(predictions_probs, book_id = 900, n = 5)
print(top_n_labels)
```

## Get full set of labels from threshold

```{r}
pred_long_all <- get_label_probabilities(predictions_probs)
```

## find tags that should be added

```{r}
add_threshold <- 0.75

final_addition_recommendations <- get_recommended_add(eav, pred_long_all, add_threshold)
```


## find tags that should be deleted

```{r}
remove_threshold <- 0.01

final_removal_recommendations <- get_recommended_remove(eav, pred_long_all, remove_threshold)
```

## given a tag, what books should it be in?

```{r}
tag_name <- "@Shaelyn Bagley"

final_single_tag_additions <- suggest_tag_additions(eav, pred_long_all, tag_name, prob_min = 0.5)
print(final_single_tag_additions)
```

## Get recommendations for a specific book

```{r}
# Get a sample book ID from the dataset
sample_book_id <- 17954


# Get recommendations for this book
recommendations <- recommend_tags_for_book(eav, predictions_probs, book_id = sample_book_id)

print(recommendations)
```
# Spectral Analysis

Goals:
1. Recommend existing tags for a set of books (and those that should be removed)
2. Discover book groupings and recommend tags that describe them.

Process for #1:

1. Form a graph based on existing tags
  a. build the stars (connect books with the same tag)
  b. connect star centers to a random other star centers for different tags (can experiment with this)
2. Spectral hashing to get a perfect code book for existing tag graph
3. Use k binary classifiers to learn the codebook (xgboost, etc.) (maps any book to codebook space)
4. Encode a book and find nearest neighbors - use those to indicate what tags should be
  a. Assume the book has the same tags as its nearest neighbor
  b. Use distance to other clusters for multi-tags

Process for #2:

1. Using encoders from step 1 encode all unseen books
2. Scan through all of possible clustering and see what's grouped (silhouette coeff?)
  a. One will likely hit the way we've tagged with unseen or misclassified books as noise
  b. Other clusters will look good (subsets, supersets, or other)
3. Find representative features for the clusters (chi-Sq or mutual information)


```{r}
tag_membership <- eav %>% 
  filter(feature == "tag") %>% 
  mutate(tag = value) %>% 
  select(id, tag) %>% 
  group_by(tag) %>% 
  summarize(membership = list(id), .groups = "drop")

# TODO: produce a list of pairs encoding a star graph over these tags. Pick a random book id from each membership list and produce a pair connecting that book id to every other book id in the membership list. This needs to be symmetric, when emitting a pair [a, b] also emit [b, a]

# Step 1: Create star graph + record center nodes per tag
star_graph_df <- tag_membership %>%
  mutate(center = purrr::map_int(membership, ~ sample(.x, 1))) %>%
  mutate(pairs = purrr::map2(membership, center, function(ids, center) {
    others <- setdiff(ids, center)
    if (length(others) == 0) return(tibble(from = integer(), to = integer()))
    bind_rows(
      tibble(from = center, to = others),
      tibble(from = others, to = center)
    )
  }))

# Step 2: Collect positive edges
positive_edges <- star_graph_df %>%
  select(pairs) %>%
  tidyr::unnest(pairs) %>%
  mutate(weight = 1)

# Step 3: Create negative edges by pairing each center with another random center
centers <- star_graph_df$center
n_centers <- length(centers)

negative_edges <- tibble(
  from = centers,
  to = purrr::map_int(centers, function(center) {
    sample(setdiff(centers, center), 1)
  })
) %>%
  bind_rows(transmute(., from = to, to = from)) %>%  # symmetric
  mutate(weight = -1)

# Step 4: Combine
star_graph_pairs <- bind_rows(positive_edges, negative_edges)

```
## Laplacian
```{r}

# Use signed weights directly
edge_list <- star_graph_pairs %>%
  select(from, to, weight)

# Step 1: Build node index
nodes <- sort(unique(c(edge_list$from, edge_list$to)))
node_index <- setNames(seq_along(nodes), nodes)

# Step 2: Build signed adjacency matrix A
A <- Matrix::sparseMatrix(
  i = node_index[as.character(edge_list$from)],
  j = node_index[as.character(edge_list$to)],
  x = edge_list$weight,
  dims = c(length(nodes), length(nodes)),
  dimnames = list(nodes, nodes)
)

# Step 3: Compute signed degree matrix D (sum of absolute edge weights)
d_vals <- Matrix::rowSums(abs(A))
D_inv_sqrt <- Matrix::Diagonal(x = 1 / sqrt(d_vals))

# Step 4: Compute signed normalized Laplacian
L_signed_norm <- Matrix::Diagonal(n = length(nodes)) - D_inv_sqrt %*% A %*% D_inv_sqrt

```
## Eigenvectors and values
```{r}
# Set number of nontrivial eigenpairs desired
k <- 5  # example value

eig <- RSpectra::eigs_sym(
  L_signed_norm,
  k       = k + 1,  # request one extra to drop trivial eigenvalue
  which   = "SM",
  target  = 0
)

# Remove the trivial eigenpair (usually eigenvalue ~0)
# Optionally filter based on a tolerance if needed
eigvals <- eig$values[-1]
eigvecs <- eig$vectors[, -1]

eigvals

```
## Iterative quantization
```{r}
iterative_quantization <- function(X, n_iter = 50) {
  # Step 1: Zero-center
  X_centered <- scale(X, center = TRUE, scale = FALSE)

  # Step 2: Optional PCA (skip if X already low-dimensional)
  # SVD of centered matrix
  svd_X <- svd(X_centered)
  V <- svd_X$v  # principal directions

  # Step 3: Project X to PCA space
  X_pca <- X_centered %*% V

  # Step 4: Initialize rotation matrix R as identity
  R <- diag(ncol(X))

  for (i in 1:n_iter) {
    # Step 5: Compute binary code
    B <- sign(X_pca %*% R)

    # Step 6: Solve for optimal R using SVD
    C <- t(B) %*% X_pca
    svd_C <- svd(C)
    R <- svd_C$u %*% t(svd_C$v)
  }

  # Final binary codes and rotation
  list(binary_code = sign(X_pca %*% R), rotation = R)
}

# eigvecs: matrix of size n_books × k from Laplacian
itq_result <- iterative_quantization(eigvecs)

binary_embedding <- itq_result$binary_code  # contains -1/+1 codes per book
rotation_matrix <- itq_result$rotation

```


